---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 12, fig.height=7,
  comment = "#>",
  fig.path = "man/figures/"
)

require(tidyverse,quietly=TRUE)
require(party,quietly=TRUE)
require(randomForestSRC,quietly=TRUE)
require(pec,quietly=TRUE)
require(survival,quietly=TRUE)
require(CoxBoost,quietly=TRUE)
require(obliqueRSF,quietly=TRUE)
require(ggthemes, quietly=TRUE)

nboots=5

```



# Overview

Oblique random survival forest (ORSFs) are ensembles for right-censured survival data that use linear combinations of input variables to recursively partition a set of training data. Regularized Cox proportional hazard models identify optimal linear combinations of input variables in each recursive partitioning step while building survival trees.

# Installation

You can install obliqueRSF from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("bcjaeger/obliqueRSF")
```

# Usage

The `ORSF` function is the center piece of the `obliqueRSF` package 


```{r usage, echo=TRUE,fig.width=10,fig.height=6}
data("pbc",package='survival')

# event is death
# censor study participants at time of last contact or transplant
pbc$status[pbc$status>=1]=pbc$status[pbc$status>=1]-1

# format categorical variables as factors
# if we don't do this, missforest will not impute 0/1 variables 
# in the way that we would like it to.
pbc = pbc %>% 
  dplyr::select(-id)%>%
  mutate(trt=factor(trt),
         ascites=factor(ascites),
         hepato=factor(hepato),
         spiders=factor(spiders),
         edema=factor(edema),
         stage=factor(stage,ordered=TRUE),
         time=time/365.25) 

orsf=ORSF(data=pbc, # data to fit trees with
          ntree=200, # number of trees to fit
          eval_times=c(1:10), # when will predictions be made?
          # note: eval_times will be used to make figures
          verbose=F) # suppresses console output
```

The `vdplot` function allows you to quickly look at the patterns in the predictions from an ORSF.

```{r vdplot1}
# Variable dependence plot (vdplot)

# Survival probabilities for a continuous variable
# note the use of sub_times, which allows you to pick
# one or more times in the evaluation times of an ORSF object

vdplot(object=orsf, xvar='bili', xlab='Bilirubin levels', 
       xvar_units = 'mg/dl', sub_times = 5)
```

here is another application of `vdplot` with a continuous x-variable, but this time we will show predictions at three times: 1 year, 3 years, and 5 years since baseline.

```{r vdplot2}
vdplot(object=orsf, xvar='albumin', xlab='Serum albumin', 
       xvar_units = 'g/dl', sub_times = c(1,3,5))
```

The `vdplot` function also supports categorical x-variables. Setting the x-variable to `sex`, and the facet variable to `hepato`,  we find a clear interaction between sex and the presence of an enlarged liver (i.e., `hepato=1`).

```{r vdplot3}
vdplot(object=orsf, xvar='sex', xlab=c("Sex"), xlvls=c("Male","Female"),
       fvar='hepato',flab=c("Normal size liver","Enlarged liver"))
```

Is this interaction something that is purely explained by sex, or is it a confounding effect from other variables that are not the same between the two sexes? We can address this using the partial dependence plot (`pdplot`) function:

```{r pdplot1, fig.width=16}
pdplot(object=orsf, xvar='sex',xlab='Sex', xlvls=c("Male","Female"),
       fvar='hepato',flvls=c("Normal size liver","Enlarged liver"),
       sub_times=c(1,3,5,7,9))
```

Taking into account the effects of other variables in the data, the interaction between sex and hepato is attenuated. 

# Performance

`orsf` objects take a long time to grow, but they usually provide excellent predictions. Below we use the `pec` package to compare the integrated Brier scores over `r nboots` replications of bootstrap cross validation using the complete cases in the `pbc` data set.

```{r performance}

# it takes a bit more code to incorporate imputation 
# into bootstrap cross-validation. However, results
# are similar using only the complete observations,
# so that's what we will use here

pbc_cmp=na.omit(pbc)

ntree=150
nsize=20

mdls=list(
  orsf=ORSF(data=pbc_cmp, ntree=ntree, use.cv = FALSE, 
            min_obs_to_split_node = nsize, verbose=FALSE),
  rsf =rfsrc(Surv(time,status)~.,data=pbc_cmp,
             ntree=ntree,nodesize=nsize),
  bws = pec::selectCox(Surv(time,status)~., data=pbc_cmp),
  cboost = pec::coxboost(Hist(time,status)~.,data=pbc_cmp)
)

# for reproducibility
set.seed(32989) 

bri_score = pec::pec(mdls, data=pbc_cmp, cens.model = 'cox',
                     formula = Surv(time,status)~1, 
                     splitMethod = 'BootCv', B=nboots)

print(bri_score)


cnc_index = pec::cindex(mdls, data=pbc_cmp, cens.model = 'cox',
                        formula = Surv(time,status)~1, 
                        splitMethod = 'BootCv', B=nboots)

print(cnc_index)


```



