---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 12, fig.height=7,
  comment = "#>",
  fig.path = "man/figures/"
)

require(tidyverse,quietly=TRUE)
require(party,quietly=TRUE)
require(randomForestSRC,quietly=TRUE)
require(pec,quietly=TRUE)
require(survival,quietly=TRUE)
require(CoxBoost,quietly=TRUE)
require(obliqueRSF,quietly=TRUE)
require(ggthemes, quietly=TRUE)

nboots=5

```



# Overview

Oblique random survival forest (ORSFs) are ensembles for right-censured survival data that use linear combinations of input variables to recursively partition a set of training data. Regularized Cox proportional hazard models identify optimal linear combinations of input variables in each recursive partitioning step while building survival trees.

# Installation

You can install obliqueRSF from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("bcjaeger/obliqueRSF")
```

# Usage

The `ORSF` function is the center piece of the `obliqueRSF` package 


```{r usage, echo=TRUE,fig.width=10,fig.height=6}

data("pbc",package='survival')

# event is death
# censor study participants at time of last contact or transplant
pbc$status[pbc$status>=1]=pbc$status[pbc$status>=1]-1

# format categorical variables as factors
# if we don't do this, missforest will not impute 0/1 variables 
# in the way that we would like it to.
pbc = pbc %>% 
  dplyr::select(-id)%>%
  mutate(
    trt=factor(trt),
    ascites=factor(ascites),
    hepato=factor(hepato),
    spiders=factor(spiders),
    edema=factor(edema),
    stage=factor(stage,ordered=TRUE),
    time=time/365.25
  ) 

orsf <- ORSF(
  data=pbc, # data to fit trees with
  ntree=100, # number of trees to fit
  eval_times=c(1:10), # when will predictions be made?
  # note: eval_times will be used to make figures
  verbose=T,# suppresses console output
  compute_oob_predictions = TRUE # return OOB preds
) 

```

The `vdplot` function allows you to quickly look at the patterns in the predictions from an ORSF.

```{r vdplot1}
# Variable dependence plot (vdplot)

# Survival probabilities for a continuous variable
# note the use of sub_times, which allows you to pick
# one or more times in the evaluation times of an ORSF object

vdplot(object=orsf, xvar='bili', xlab='Bilirubin levels', 
       xvar_units = 'mg/dl', sub_times = 5)
```

here is another application of `vdplot` with a continuous x-variable, but this time we will show predictions at three times: 1 year, 3 years, and 5 years since baseline.

```{r vdplot2}
vdplot(object=orsf, xvar='albumin', xlab='Serum albumin', 
       xvar_units = 'g/dl', sub_times = c(1,3,5))
```

The `vdplot` function also supports categorical x-variables. Setting the x-variable to `sex`, and the facet variable to `hepato`,  we find a clear interaction between sex and the presence of an enlarged liver (i.e., `hepato=1`).

```{r vdplot3}
vdplot(object=orsf, xvar='sex', xlab=c("Sex"), xlvls=c("Male","Female"),
       fvar='hepato',flab=c("Normal size liver","Enlarged liver"))
```

Is this interaction something that is purely explained by sex, or is it a confounding effect from other variables that are not the same between the two sexes? We can address this using the partial dependence plot (`pdplot`) function:

```{r pdplot1, fig.width=16}
pdplot(object=orsf, xvar='sex',xlab='Sex', xlvls=c("Male","Female"),
       fvar='hepato',flvls=c("Normal size liver","Enlarged liver"),
       sub_times=c(1,3,5,7,9))
```

Taking into account the effects of other variables in the data, the interaction between sex and hepato is attenuated. 

